---
title: "Central India Connectivity Synthesis"
author: "Jay Schoen"
editor_options:
  chunk_output_type: console
---

Notes
  - Layers projected to UTM 44N
  
  - Layers resampled (bilinear interpolation) to 250m resolution
  
  - Layers masked to union shape
  
  - Yumnam nightlight areas set to 1 for resistance analysis
  

```{r}
library(sp)
library(sf)
library(raster)
library(tidyverse)
library(rgeos)
library(rgdal)
library(scales)
library(ggsci)
library(ggthemes)
```

# Setup
```{r}
# World Mercator projection
merc <- "+init=epsg:3395"

# All layers union shape
tiger_lyrs_shp <- readOGR("spatial_layers/tiger_layers_shape.shp") 

# PA shapefile
PA_shp <- readOGR("Buffer node analysis/PA_node.shp") %>%
  spTransform(merc) 

# Template raster
template250 <- raster(resolution = c(250, 250), ext = extent(tiger_lyrs_shp), crs = merc)
# template1k <- raster(resolution = c(1000, 1000), ext = extent(tiger_lyrs_shp), crs = merc)
```

# Functions used 
```{r}
## rasterRescale (from http://www.timassal.com/?p=859)
rasterRescale<-function(r){
  ((r-cellStats(r,"min"))/(cellStats(r,"max")-cellStats(r,"min")))
}

## Mode functions from Lucas Fortini (https://stackoverflow.com/questions/26726028/how-to-pick-the-most-frequent-values-mode-from-a-raster-stack)
Mode <- function(x) {
  ux <- unique(x)
  ux=ux[!is.na(ux)]
  ux[which.max(tabulate(match(x, ux)))]
}
Mode_count <- function(x) {
  ux <- unique(x)    
  ux=ux[!is.na(ux)]
  mode=ux[which.max(tabulate(match(x, ux)))]
  sum(x==mode, na.rm=T)
}
```

# Importing rasters/adjusting Yumnam
```{r}
res_list <- lapply(list.files("analysis_final/res/", pattern = "res.tif", 
                       all.files = TRUE, full.names = TRUE), raster)

stack_res <- stack(res_list) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(stack_res)

# Modifying Yumnam layer for Resistance analysis
## (setting nightlight areas (currently masked) == 1)
yumnam.raw <- raster("DATA/JHALA/Resistance1/Resistance/resist_final/w001001.adf") 

yumnam_res_nl1 <- yumnam.raw 
yumnam_res_nl1[is.na(yumnam_res_nl1[])] <- 1 # Replacing NAs (nightlights) in Yumnam raster with 1
yumnam_res_nl1 <- yumnam_res_nl1 %>%
  projectRaster(template250) 
yumnam_res_nl1[yumnam_res_nl1 >1] <- NA # removing erroneous >1 values left over from interpolation
yumnam_res_nl1[yumnam_res_nl1 <=0] <- 1e-7 # recoding negative areas to 1e-7 (lowest resistance)

stack_res[[5]] <- yumnam_res_nl1 %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp) %>%
  `names<-` ("Yumnam_res")
plot(stack_res)                               # Check: no holes in Yumnam layer
# Visualizing distributions of rasters
par(mfrow = c(1,5))
hist(stack_res, xlab = "Resistance", freq = F)
par(mfrow = c(1,1), mar = c(5,4,4,2))
```

# Scale & Average rasters
```{r}
stack_res.scaled <- rasterRescale(stack_res) %>%
  `names<-` (str_glue("{names(stack_res)}_scaled"))
plot(stack_res.scaled)
hist(stack_res.scaled)                        # Check: histograms unchanged
par(mfrow = c(1,1), mar = c(5,4,4,2))

## Mean Resistance Layer
res.scaled_avg <- mean(stack_res.scaled)
plot(res.scaled_avg, main = "Average Scaled")
hist(res.scaled_avg, main = "Scaled Average", xlab = "Avg Resistance", breaks = 30)
# writeRaster(res.scaled_avg, "Buffer node analysis/Resistance/Agreement/res.scaled_avg.tif")

## Coefficient of Variation of Mean layer
res.scaled_cv <- calc(stack_res.scaled, cv)
plot(res.scaled_cv, main = "CV of Avg")
hist(res.scaled_cv, main = "CV of Avg", xlab = "CV of Avg Resistance")
# writeRaster(res.scaled_cv, "Buffer node analysis/Resistance/Agreement/res.scaled_cv.tif")


## CV shows agreement between layers in high resistance areas, but variation in low resistance

## Overall, very high variation in values between raster layers; need to discretize to draw meaningful conclusions
```

# Discretizing (5 levels)
```{r}
quant5_res <- t(quantile(stack_res, probs = seq(0,1,.2))) # quintiles

# reclassifying values for cut function (need discrete cutoffs/lower min bound to include minimum values)
quant5_res[1,] <- c(0.999, 0.0838757, 24.99167, 0.999, 0) ; 
quant5_res[3,1] <- 100.999999
quant5_res[c(2,4),4] <- c(1.000001, 4.999999)
quant5_res[5,5] <- 0.999

quant5_res[c(1,4),4] <- c(0.999, 4.999999)
quant5_res

stack_res.disc5 = stack(lapply(names(stack_res), function(n)
  cut(stack_res[[n]], breaks=quant5_res[,n])))
names(stack_res.disc5) <- str_glue("{names(stack_res)}_discrete5")
plot(stack_res.disc5)

par(mfrow = c(1,5))
hist(stack_res.disc5, breaks = c(0:5)) # irregularities due to extremely skewed distributions
par(mfrow = c(1,1), mar = c(5,4,4,2))


## Export
# writeRaster(stack_res.disc5, file.path("analysis_final/res/", names(stack_res.disc5)), bylayer = T, format = "GTiff", overwrite = T)

## Import
res.disc5_list <- lapply(list.files("analysis_final/res/",
                                    pattern = "discrete5.tif", all.files = TRUE, 
                                    full.names = TRUE), raster)

stack_res.disc5 <- stack(res.disc5_list) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(stack_res.disc5)
```

## Mode analysis
```{r}
res_disc5.mode <- calc(stack_res.disc5, fun=Mode)
plot(res_disc5.mode, main = "Mode Value")
# writeRaster(res_disc5.mode, "analysis_final/res/res_disc5.mode.tif", overwrite = T)
# 
# res_disc5.sd <- calc(stack_res.disc5, sd)
# plot(res_disc5.sd, main = "SD of Mode Value")
# # writeRaster(res_disc5.sd, "analysis_final/res/res_disc5.sd.tif", overwrite = T)

res_disc5.mode.freq <- calc(stack_res.disc5, fun=Mode_count) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(res_disc5.mode.freq, main = "Frequency of mode value")

res_disc5.perc.agree <- ((res_disc5.mode.freq/5) * 100) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(res_disc5.perc.agree, main = "% Agreement on mode value")
# writeRaster(res_disc5.perc.agree, "analysis_final/res/res_disc5.perc.agree.tif", overwrite = T)


## Import
# res_disc5.mode <- raster("analysis_final/res/res_disc5.mode.tif")
# res_disc5.perc.agree <- raster("analysis_final/res/res_disc5.perc.agree.tif")
```

#### Agreement/Quantile plots
```{r}
# Quantiles + Agreement plot
res_disc5.mode_data <- data.frame(quantile = as.factor(getValues(res_disc5.mode))) %>%
  mutate(perc.agree = as.factor(getValues(res_disc5.perc.agree))) %>%
  drop_na()

plot_res_disc5.agree <- ggplot(data = res_disc5.mode_data, 
                             aes(x = perc.agree, fill = quantile)) +
  geom_bar(position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = c("steelblue", "green4", "yellow3", "orange3", "red3"), 
                    labels = c("1: Low", "2: Med-Low", "3: Medium", "4: Med-High", "5: High"),
                    name = "Resistance Quantile") +
  # theme_bw() + 
  theme(panel.background = element_rect(fill = NA, color = NA),
        panel.grid = element_line(color = "gray90")) +
  # ggtitle("Resistance % Agreement by Quantile") +
  ylab("Number of pixels") +
  xlab("Agreement (%)")

plot(plot_res_disc5.agree)

plot2_res_disc5.agree <- ggplot(data = res_disc5.mode_data, 
                             aes(x = quantile, fill = perc.agree)) +
  geom_bar(position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = c("red3", "orange2", "darkolivegreen4", "skyblue2", "blue4"),
                    name = "% Agreement") +
  theme_bw() + 
  ggtitle("Resistance % Agreement by Quantile") +
  ylab("# Pixels") +
  xlab("Quantile")

plot(plot2_res_disc5.agree)


# ggsave(plot_res_disc5.agree, filename = "analysis_final/figures/fig3a.svg", width = 7, height = 4, dpi = 300)
# ggsave(plot2_res_disc5.agree, filename = "analysis_final/res/plot2_res_disc5_agree.png", width = 7, height = 4, dpi = 300)
```

## Land Cover Analysis
```{r}
lc <- raster("spatial_layers/vegetation_typemap.img")
# lc_250 <- projectRaster(lc, template250, method = "ngb") 
# writeRaster(lc, "spatial_layers/vegetation_typemap.tif", overwrite = T)
# writeRaster(lc_250, "spatial_layers/lc_250.tif", overwrite = T)

lc_250 <- raster("spatial_layers/lc_250.tif")
plot(lc_250)

lc_data <- data.frame(lc@data@attributes) %>%
  select(ID, COUNT, Class_Names) %>%
  filter(COUNT>0)

lc_250_data <- data.frame(ID = getValues(lc_250), 
                          perc.agree = as.factor(getValues(res_disc5.perc.agree)),
                          mode.quant = as.factor(getValues(res_disc5.mode))) %>%
  drop_na()

lc_top11 <- lc_250_data %>%             # 11/255 classes are at least 1% of total pixels
  group_by(ID) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:11) %>%
  left_join(lc_data, by = "ID") %>%
  # filter(Class_Names != "Agriculture") %>%
  select(-COUNT) %>%
  mutate(prop = count/sum(count),
         Class_Names = factor(Class_Names, levels = .$Class_Names))

plot_lc_top11 <- ggplot(lc_top11, aes(x = Class_Names, y = prop, fill = Class_Names)) +
  geom_col() + 
  scale_fill_tableau() +
  theme_bw() +
  ggtitle("Proportion of Landcover") +
  xlab("Landcover Class") +
  ylab("Proportion") +
  scale_x_discrete(label=function(x) abbreviate(x, minlength=15)) +
  theme(axis.text.x = element_text(angle = 15, vjust = .7))

plot_lc_top11
```

### LC/Agreement plot
```{r}
res_lc_data1 <- lc_250_data %>%
  left_join(lc_top11, by = "ID") %>%
  filter(ID %in% lc_top11$ID)

res_lc_data2 <- res_lc_data1 %>%
  select(-count, -prop) %>%
  mutate(count = 1) %>%
  # filter(Class_Names != "Agriculture") %>%
  group_by(perc.agree, Class_Names) %>%
  summarise(count = sum(count)) %>%
  mutate(prop.perc = count/sum(count)) %>%
  ungroup() %>%
  group_by(Class_Names) %>%
  mutate(prop.class = count/sum(count)) %>%
  ungroup() %>%
  mutate(Class_Names = factor(Class_Names, levels = lc_top11$Class_Names))
```

#### Aggregating classes
```{r}
lc_agg <- res_lc_data1 %>%
  mutate(Class_Names = recode(Class_Names, 
                              'Sal' = "Forest",
                              'Dry deciduous' = "Forest",
                              'Teak mixed moist deciduous' = "Forest",
                              'Moist deciduous' = "Forest",
                              'Teak mixed dry deciduous' = "Forest",
                              'Teak' = "Forest")) %>%
  # filter(Class_Names != "Agriculture") %>%            ### to look at LC besides agriculture
  select(-count, -prop) %>%
  mutate(count = 1) %>%
  group_by(perc.agree, Class_Names) %>%
  summarise(count = sum(as.numeric(count))) %>%
  group_by(perc.agree) %>%
  mutate(prop.perc = count/sum(count)) %>%
  ungroup() %>%
  group_by(Class_Names) %>%
  mutate(prop.class = count/sum(count))

plot2_res_lc_agg.agree <- ggplot(data = lc_agg, 
                             aes(x = perc.agree, y = prop.perc, fill = Class_Names)) +
  geom_col(position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = c("grey50", "dark green", "yellow4", "orange4", "steel blue4", "red4"),
                    name = "Landcover") +
  # theme_bw() + 
  theme(panel.background = element_rect(fill = NA, color = NA),
        panel.grid = element_line(color = "gray90")) +
  # ggtitle("Resistance % Agreement by LC") +
  ylab("Proportion of pixels") +
  xlab("Agreement (%)")

plot(plot2_res_lc_agg.agree)

# ggsave(plot2_res_lc_agg.agree, filename = "analysis_final/figures/fig3b.svg", width = 7, height = 4, dpi = 300)

```



## Agreement Layers
```{r}
# 60% or more agreement (High agreement)
res_mask60 <- res_disc5.perc.agree %in% 60:100
res_disc5.mode_agree.60 <- mask(res_disc5.mode, res_mask60, maskvalue = 0)
plot(res_disc5.mode_agree.60)

## Low resistance in >60% agreement
low.res1_disc5_agree.60 <- res_disc5.mode_agree.60 %in% 1 %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
  
low.res1_disc5_agree.60[low.res1_disc5_agree.60 !=1] <- NA
plot(low.res1_disc5_agree.60)

# writeRaster(low.res1_disc5_agree.60, "analysis_final/res/res_disc5_low1_agree.60.tif", overwrite = T)

## High resistance in >60% agreement
high.res1_disc5_agree.60 <- res_disc5.mode_agree.60 %in% 5 %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)

high.res1_disc5_agree.60[high.res1_disc5_agree.60 !=1] <- NA
plot(high.res1_disc5_agree.60)

# writeRaster(high.res1_disc5_agree.60, "analysis_final/res/res_disc5_high1_agree.60.tif", overwrite = T)


# 80% or more agreement (Very high agreement)
res_mask80 <- res_disc5.perc.agree %in% 80:100
res_disc5.mode_agree.80 <- mask(res_disc5.mode, res_mask80, maskvalue = 0)
plot(res_disc5.mode_agree.80)
# writeRaster(res_disc5.mode_agree.80, "analysis_final/res/res_disc5_mode_agree_80.tif", overwrite = T)


## In QGIS, to create sieved shape file:
      # imported low.res_agree file
      # sieved at 10
      # polygonized
      # clipped to tiger_lyrs_shp
      # selected all features == 1 (focal areas)
      # saved as new shape file (low.res1_disc5_agree.60_sieve10.shp)
```

## Quantitative Analysis
```{r}
# High agreement summary table
high_agree60 <- res_disc5.mode_data %>%
  mutate(perc.agree = as.numeric(perc.agree)) %>%
  filter(perc.agree >= 3) %>%
  nrow() / nrow(res_disc5.mode_data)

# Very high agreement summary table
high_agree80 <- res_disc5.mode_data %>%
  mutate(perc.agree = as.numeric(perc.agree)) %>%
  filter(perc.agree >= 4) %>%
  nrow() / nrow(res_disc5.mode_data)

# Agreement by quantile table
agree_by_quant <- res_disc5.mode_data %>% 
  group_by(perc.agree, quantile) %>%
  summarize(perc_quant = n()) %>%
  group_by(perc.agree) %>%
  mutate(prop = perc_quant/sum(perc_quant))
```


# Discretizing into 3,4,6 levels (for Appendix)
```{r}
quant4_res <- t(quantile(stack_res, probs = seq(0,1,.25))) # quartiles

# reclassifying values for cut function (lower min bound to include minimum values)
quant4_res[1,] <- c(0.999, 0.0838757, 24.99167, 0.999, 0)

stack_res.disc4 = stack(lapply(names(stack_res), function(n)
  cut(stack_res[[n]], breaks=quant4_res[,n])))
names(stack_res.disc4) <- str_glue("{names(stack_res)}_discrete4")
plot(stack_res.disc4)

par(mfrow = c(1,5))
hist(stack_res.disc4, breaks = c(0:4)) # irregularities due to extremely skewed distributions
par(mfrow = c(1,1), mar = c(5,4,4,2))


## Export
# writeRaster(stack_res.disc4, file.path("analysis_final/res/", names(stack_res.disc4)), bylayer = T, format = "GTiff", overwrite = T)
```

```{r}
quant3_res <- t(quantile(stack_res, probs = seq(0,1,(1/3)))) # three levels

# reclassifying values for cut function (need discrete cutoffs/lower minimums)
quant3_res[1,] <- c(0.999, 0.0838757, 24.99167, 0.999, 0) ; 


stack_res.disc3 = stack(lapply(names(stack_res), function(n)
  cut(stack_res[[n]], breaks=quant3_res[,n])))
names(stack_res.disc3) <- str_glue("{names(stack_res)}_discrete3")
plot(stack_res.disc3)

par(mfrow = c(1,5))
hist(stack_res.disc3, breaks = c(0:3)) # irregularities due to extremely skewed distributions
par(mfrow = c(1,1), mar = c(5,4,4,2))


## Export
# writeRaster(stack_res.disc3, file.path("analysis_final/res/", names(stack_res.disc3)), bylayer = T, format = "GTiff", overwrite = T)
```


```{r}
quant6_res <- t(quantile(stack_res, probs = seq(0,1,(1/6)))) # six levels

# reclassifying values for cut function (need discrete cutoffs/lower minimums)
quant6_res[1,] <- c(0.999, 0.0838757, 24.99167, 0.999, 0) ; 
quant6_res [5,c(4,5)] <- c(4.99999, 0.999)
quant6_res[5,5] <- 0.999


stack_res.disc6 = stack(lapply(names(stack_res), function(n)
  cut(stack_res[[n]], breaks=quant6_res[,n])))
names(stack_res.disc6) <- str_glue("{names(stack_res)}_discrete6")
plot(stack_res.disc6)

par(mfrow = c(1,5))
hist(stack_res.disc6, breaks = c(0:6)) # irregularities due to extremely skewed distributions
par(mfrow = c(1,1), mar = c(5,4,4,2))


## Export
# writeRaster(stack_res.disc6, file.path("analysis_final/res/", names(stack_res.disc6)), bylayer = T, format = "GTiff", overwrite = T)
```
