---
title: "Central India Connectivity Synthesis"
author: "Jay Schoen"
editor_options: 
  chunk_output_type: console
---

Notes

  - Current analysis performed in Circuitscape 4.0 standalone GUI
    	- Resistance rasters with raw values (after alignment and 250m projection)
    	- PA node layer from Prachi + Kanha buffer + Panna buffer (no PA buffers besides Panna/Kanha) as node layer
    	- Pairwise modeling mode
	    - Cumulative pairwise curent flow option selected
	    - Nodes left hot (not set to 0)

```{r}
library(sp)
library(sf)
library(raster)
library(tidyverse)
library(rgeos)
library(rgdal)
library(scales)
```

# Setup
```{r}
utm.44n <- "+init=epsg:3395"

# All layers union shape
tiger_lyrs_shp <- readOGR("spatial_layers/tiger_layers_shape.shp") 

# PA shapefile
PA_shp <- readOGR("Buffer node analysis/PA_node.shp") %>%
  spTransform(utm.44n) 

# Template raster
template250 <- raster(resolution = c(250, 250), ext = extent(tiger_lyrs_shp), crs = utm.44n)
# template1k <- raster(resolution = c(1000, 1000), ext = extent(tiger_lyrs_shp), crs = utm.44n)
```

# Functions used 
```{r}
## rasterRescale (from http://www.timassal.com/?p=859)
rasterRescale<-function(r){
  ((r-cellStats(r,"min"))/(cellStats(r,"max")-cellStats(r,"min")))
}

## Mode functions from Lucas Fortini (https://stackoverflow.com/questions/26726028/how-to-pick-the-most-frequent-values-mode-from-a-raster-stack)
Mode <- function(x) {
  ux <- unique(x)
  ux=ux[!is.na(ux)]
  ux[which.max(tabulate(match(x, ux)))]
}
Mode_count <- function(x) {
  ux <- unique(x)    
  ux=ux[!is.na(ux)]
  mode=ux[which.max(tabulate(match(x, ux)))]
  sum(x==mode, na.rm=T)
}
```



# Notes
  - All layers run in Circuitscape standalone GUI wth PA_node as node layer
  - Nodes left hot (not set to 0), cumulative pairwise curent flow option selected
  - Results imported below

# Importing CS results
```{r}
## Masking CS results outside of individual study borders
dutta_cur.raw <- raster("C:/CS_dutta250m_PA_node/CS_dutta250m_PA_node_cum_curmap.asc", 
                        crs = utm.44n)
mondal_cur.raw <- raster("C:/CS_mondal250m_PA_node/CS_mondal250m_PA_node_cum_curmap.asc", 
                         crs = utm.44n)
reddy_cur.raw <- raster("C:/CS_reddy250m_PA_node/CS_reddy250m_PA_node_cum_curmap.asc", 
                        crs = utm.44n)
thatte_cur.raw <- raster("C:/CS_thatte250m_PA_node/CS_thatte250m_PA_node_cum_curmap.asc", 
                         crs = utm.44n)
yumnam_cur.raw <- raster("C:/CS_yumnam250m_PA_node_nl/CS_yumnam250m_PA_node_nl_cum_curmap.asc", 
                         crs = utm.44n)

stack_cur.raw <- stack(dutta_cur.raw, mondal_cur.raw, reddy_cur.raw, 
                       thatte_cur.raw, yumnam_cur.raw) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp) %>%
  `names<-` (str_glue("{lyr_names}_cur"))

plot(stack_cur.raw)
hist(stack_cur.raw)
par(mfrow = c(1,1), mar = c(5,4,4,2))

stack_cur.raw[stack_cur.raw > 12] <- NA # masking out nodes (only node areas were over 12)
plot(stack_cur.raw)
hist(stack_cur.raw)
par(mfrow = c(1,1), mar = c(5,4,4,2))

# writeRaster(stack_cur.raw, file.path("analysis_final/cur/", names(stack_cur.raw)), bylayer = T, format = "GTiff", overwrite = T)


# Importing TIFF Rasters 
cur_list <- lapply(list.files("analysis_final/cur/", pattern = "_cur.tif", 
                              all.files = TRUE, full.names = TRUE, ), raster)

stack_cur <- stack(cur_list) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(stack_cur)
hist(stack_cur, xlab = "Current")
par(mfrow = c(1,1), mar = c(5,4,4,2))
```

## Scale/average
```{r}
stack_cur.scaled <- rasterRescale(stack_cur) %>%
  `names<-` (str_glue("{names(stack_cur)}_scaled"))
plot(stack_cur.scaled)
hist(stack_cur.scaled)
par(mfrow = c(1,1), mar = c(5,4,4,2))

cur.scaled_avg <- mean(stack_cur.scaled)
plot(cur.scaled_avg, main = "Average Current (Scaled)")
hist(cur.scaled_avg, main = "Scaled Average Current", xlab = "Current", breaks = 30, xlim = c(0,0.5))

cur.scaled_cv <- calc(stack_cur.scaled, cv)
plot(cur.scaled_cv, main = "CV of Avg Current (Scaled)")
hist(cur.scaled_cv, main = "Scaled CV Current")

## Exporting files for GIS
# writeRaster(cur.scaled_avg, "analysis_final/cur/cur_scaled_avg.tif", overwrite = TRUE)
# writeRaster(cur.scaled_cv, "analysis_final/cur/cur_scaled_cv.tif", overwrite = TRUE)

## Import
cur.scaled_avg <- raster("analysis_final/cur/cur_scaled_avg.tif")
cur.scaled_cv <- raster("analysis_final/cur/cur_scaled_cv.tif")
```

### 20%ile CV
```{r}
  # Low 20%ile CV areas
cur.cv_quant5 <- quantile(cur.scaled_cv, probs = seq(0,1,.2))
cur.scaled_low.cv20 <- cur.scaled_cv
cur.scaled_low.cv20[cur.scaled_low.cv20 > cur.cv_quant5[2]] <- NA
plot(cur.scaled_low.cv20)

cur.scaled_avg_low.cv20 <- mask(cur.scaled_avg, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_low.cv20)
# writeRaster(cur.scaled_avg_low.cv20, "analysis_final/cur/cur_scaled_avg_low_cv20.tif")

# Low 20%ile CV High Current 

## >50%ile of current masked with <20%ile CV
cur.avg_quant10 <- quantile(cur.scaled_avg, probs = seq(0,1,.1))
cur.scaled_avg_high.cur50 <- cur.scaled_avg
cur.scaled_avg_high.cur50[cur.scaled_avg_high.cur50 < cur.avg_quant10[6]] <- NA
plot(cur.scaled_avg_high.cur50)
cur.scaled_avg_high.cur50_low.cv20 <- mask(cur.scaled_avg_high.cur50, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur50_low.cv20)

# writeRaster(cur.scaled_avg_high.cur50, "analysis_final/cur/cur_scaled_avg_high_cur50.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur50_low.cv20, 
#             "analysis_final/cur/cur_scaled_avg_high_cur50_low_cv20.tif")

## >60%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur60 <- cur.scaled_avg
cur.scaled_avg_high.cur60[cur.scaled_avg_high.cur60 < cur.avg_quant10[7]] <- NA
plot(cur.scaled_avg_high.cur60)
cur.scaled_avg_high.cur60_low.cv20 <- mask(cur.scaled_avg_high.cur60, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur60_low.cv20)

# writeRaster(cur.scaled_avg_high.cur60, "analysis_final/cur/cur_scaled_avg_high_cur60.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur60_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur60_low_cv20.tif")

## >70%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur70 <- cur.scaled_avg
cur.scaled_avg_high.cur70[cur.scaled_avg_high.cur70 < cur.avg_quant10[8]] <- NA
plot(cur.scaled_avg_high.cur70)
cur.scaled_avg_high.cur70_low.cv20 <- mask(cur.scaled_avg_high.cur70, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur70_low.cv20)

# writeRaster(cur.scaled_avg_high.cur70, "analysis_final/cur/cur_scaled_avg_high_cur70.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur70_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur70_low_cv20.tif")

## >80%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur80 <- cur.scaled_avg
cur.scaled_avg_high.cur80[cur.scaled_avg_high.cur80 < cur.avg_quant10[9]] <- NA
plot(cur.scaled_avg_high.cur80)
cur.scaled_avg_high.cur80_low.cv20 <- mask(cur.scaled_avg_high.cur80, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur80_low.cv20)

# writeRaster(cur.scaled_avg_high.cur80, "analysis_final/cur/cur_scaled_avg_high_cur80.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur80_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur80_low_cv20.tif")

## >90%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur90 <- cur.scaled_avg
cur.scaled_avg_high.cur90[cur.scaled_avg_high.cur90 < cur.avg_quant10[10]] <- NA
plot(cur.scaled_avg_high.cur90)
cur.scaled_avg_high.cur90_low.cv20 <- mask(cur.scaled_avg_high.cur90, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur90_low.cv20)

# writeRaster(cur.scaled_avg_high.cur90, "analysis_final/cur/cur_scaled_avg_high_cur90.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur90_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur90_low_cv20.tif")
```

### 50%ile CV
```{r}
# Low 50%ile CV areas
cur.cv_quant2 <- quantile(cur.scaled_cv, probs = seq(0,1,.5))
cur.scaled_low.cv50 <- cur.scaled_cv
cur.scaled_low.cv50[cur.scaled_low.cv50 > cur.cv_quant2[2]] <- NA
plot(cur.scaled_low.cv50)

cur.scaled_avg_low.cv50 <- mask(cur.scaled_avg, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_low.cv50)
# writeRaster(cur.scaled_avg_low.cv50, "analysis_final/cur/cur_scaled_avg_low_cv50.tif")

## >60%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur60_low.cv50 <- mask(cur.scaled_avg_high.cur60, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur60_low.cv50)
# writeRaster(cur.scaled_avg_high.cur60_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur60_low_cv50.tif")

## >70%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur70_low.cv50 <- mask(cur.scaled_avg_high.cur70, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur70_low.cv50)
# writeRaster(cur.scaled_avg_high.cur70_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur70_low_cv50.tif")

## >80%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur80_low.cv50 <- mask(cur.scaled_avg_high.cur80, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur80_low.cv50)
# writeRaster(cur.scaled_avg_high.cur80_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur80_low_cv50.tif")

## >90%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur90_low.cv50 <- mask(cur.scaled_avg_high.cur90, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur90_low.cv50)
# writeRaster(cur.scaled_avg_high.cur90_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur90_low_cv50.tif")

```

# Quantitative Analysis

## Total CCA area

```{r}
cca_area <- as.numeric(length(cur.scaled_avg_high.cur70_low.cv20[!is.na(cur.scaled_avg_high.cur70_low.cv20)])) * 250*250/1000
```


## High current/high agreement overlap
```{r}
#20th %ile CV used for high agreement
plot(cur.scaled_low.cv20)

#70th %ile current
low.cv20_high.cur70_mask <- mask(cur.scaled_low.cv20, cur.scaled_avg_high.cur70)
plot(low.cv20_high.cur70_mask)

length(low.cv20_high.cur70_mask[!is.na(low.cv20_high.cur70_mask)])/
  length(cur.scaled_low.cv20[!is.na(cur.scaled_low.cv20)])


# 80th %ile current
low.cv20_high.cur80_mask <- mask(cur.scaled_low.cv20, cur.scaled_avg_high.cur80)
plot(low.cv20_high.cur80_mask)

length(low.cv20_high.cur80_mask[!is.na(low.cv20_high.cur80_mask)])/
  length(cur.scaled_low.cv20[!is.na(cur.scaled_low.cv20)])

plot(cur.scaled_avg_high.cur80)
high.cur80_low.cv20_mask <- mask(cur.scaled_avg_high.cur80, cur.scaled_low.cv20)
plot(high.cur80_low.cv20_mask)

length(high.cur80_low.cv20_mask[!is.na(high.cur80_low.cv20_mask)])/
  length(cur.scaled_avg_high.cur80[!is.na(cur.scaled_avg_high.cur80)])


int_low.cv20_high.cur80 <- hull_low.cv20 %>%
  crop(hull_high.cur80) %>%
  mask(hull_high.cur80)
plot(int_low.cv20_high.cur80)

int_high.cur80_low.cv20 <- hull_high.cur80 %>%
  crop(hull_low.cv20) %>%
  mask(hull_low.cv20)
plot(int_high.cur80_low.cv20)

length(int_low.cv20_high.cur80[!is.na(int_low.cv20_high.cur80)])/
  length(hull_low.cv20[!is.na(hull_low.cv20)])

length(int_low.cv20_high.cur80[!is.na(int_low.cv20_high.cur80)])/
  length(hull_low.cv20[!is.na(hull_low.cv20)])

```
