---
title: "Central India Connectivity Synthesis"
author: "Jay Schoen"
editor_options: 
  chunk_output_type: console
---

Notes

  - Current analysis performed in Circuitscape 4.0 standalone GUI
    	- Resistance rasters with raw values (after alignment and 250m projection)
    	- PA node layer from Prachi + Kanha buffer + Panna buffer (no PA buffers besides Panna/Kanha) as node layer
    	- Pairwise modeling mode
	    - Cumulative pairwise curent flow option selected
	    - Nodes left hot (not set to 0)
	    
	- Nodeless details
	    - 130km ring buffer (20% of study extent (650 km wide))
	    - 20 nodes randomly placed in perimeter
	    - identical Circuitscape settings as above
	    - buffer removed after Circuitscape

```{r}
library(sp)
library(sf)
library(raster)
library(tidyverse)
library(rgeos)
library(rgdal)
library(scales)
```

# Setup
```{r}
utm.44n <- "+init=epsg:3395"


# All layers union shape
tiger_lyrs_shp <- readOGR("spatial_layers/tiger_layers_shape.shp") 

# Creating ring buffer
ext_buffer <- buffer(tiger_lyrs_shp, 150000)
plot(ext_buffer); plot(tiger_lyrs_shp, col = "grey", add = T);axis(1);axis(2)

## Template raster
template250 <- raster(resolution = c(250, 250), ext = extent(ext_buffer), crs = utm.44n)

ext_buf_r <- rasterize(ext_buffer, template250)
ext_buf_r[] <- rnorm(length(getValues(ext_buf_r)), mean = .5, sd = .1)
plot(ext_buf_r)

ring_buf <- gDifference(ext_buffer, tiger_lyrs_shp) 

ring_buf_r <- ext_buf_r %>%
  crop(ring_buf) %>%
  mask(ring_buf)

plot(ring_buf, add = T);axis(1);axis(2)

plot(ring_buf_r)


# Nodes
## Imported from QGIS (easier than in R)
##  - 20 random points in ring buffer, minimum 100 km apart

ring_nodes <- readOGR("analysis_final/nodeless/ring_nodes.shp")
ring_nodes@data$objectID <- 1:20
ring_nodes_r <- rasterize(ring_nodes, ring_buf_r, field = "objectID")

ring_nodes_r[!is.na(ring_nodes_r)]
plot(ring_nodes_r)

writeRaster(ring_nodes_r, "analysis_final/nodeless/ring_nodes.asc")
```


# Creating ring buffers

Ring value rules
  - random draw from normal distribution
  - mean = median of layer
  - sd = sd of layer
  - constrained between min/max of original layer (to prevent <0 or very high values)

```{r}

## Importing/stacking aligned resistance rasters
res_list <- lapply(list.files("analysis_final/res/", pattern = "res.tif", 
                       all.files = TRUE, full.names = TRUE), raster)

stack_res <- stack(res_list) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(stack_res)


# Visualizing distributions of rasters
hist(stack_res, xlab = "Resistance", freq = F)
par(mfrow = c(1,1), mar = c(5,4,4,2))
```

## Dutta
```{r}
dutta <- stack_res[[1]]
dutta_vals <- getValues(dutta) %>% as_tibble() %>% drop_na()
dutta_vals <- dutta_vals$value
dutta_max <- max(dutta_vals)
dutta_min <- min(dutta_vals)

dutta_rand <- rasterize(ext_buffer, ext_buf_r)
dutta_rand[] <- rnorm(length(getValues(dutta_rand)), 
                      mean = median(dutta_vals),
                      sd = sd(dutta_vals))
dutta_rand[dutta_rand<dutta_min] <- dutta_min
dutta_rand[dutta_rand>dutta_max] <- dutta_max

dutta_ring <- dutta_rand %>%
  crop(ring_buf) %>%
  mask(ring_buf)
  
plot(dutta_ring)

dutta_res_nodeless <- terra::cover(dutta_ring, extend(dutta, dutta_ring))
plot(dutta_res_nodeless, main = "Dutta et al.")
```

## Mondal
```{r}
mondal <- stack_res[[2]]
mondal_vals <- getValues(mondal) %>% as_tibble() %>% drop_na()
mondal_vals <- mondal_vals$value
mondal_max <- max(mondal_vals)
mondal_min <- min(mondal_vals)

mondal_rand <- rasterize(ext_buffer, ext_buf_r)
mondal_rand[] <- rnorm(length(getValues(mondal_rand)), 
                      mean = median(mondal_vals),
                      sd = sd(mondal_vals))
mondal_rand[mondal_rand<mondal_min] <- mondal_min
mondal_rand[mondal_rand>mondal_max] <- mondal_max

mondal_ring <- mondal_rand %>%
  crop(ring_buf) %>%
  mask(ring_buf)
  
plot(mondal_ring)

mondal_res_nodeless <- terra::cover(mondal_ring, extend(mondal, mondal_ring))
plot(mondal_res_nodeless, main = "Mondal et al.")
```

## Reddy
```{r}
reddy <- stack_res[[3]]
reddy_vals <- getValues(reddy) %>% as_tibble() %>% drop_na()
reddy_vals <- reddy_vals$value
reddy_max <- max(reddy_vals)
reddy_min <- min(reddy_vals)

reddy_rand <- rasterize(ext_buffer, ext_buf_r)
reddy_rand[] <- rnorm(length(getValues(reddy_rand)), 
                      mean = median(reddy_vals),
                      sd = sd(reddy_vals))
reddy_rand[reddy_rand<reddy_min] <- reddy_min
reddy_rand[reddy_rand>reddy_max] <- reddy_max

reddy_ring <- reddy_rand %>%
  crop(ring_buf) %>%
  mask(ring_buf)
  
plot(reddy_ring)

reddy_res_nodeless <- terra::cover(reddy_ring, extend(reddy, reddy_ring))
plot(reddy_res_nodeless, main = "Reddy et al.")
```


## Thatte
```{r}
thatte <- stack_res[[4]]
thatte_vals <- getValues(thatte) %>% as_tibble() %>% drop_na()
thatte_vals <- thatte_vals$value
thatte_max <- max(thatte_vals)
thatte_min <- min(thatte_vals)

thatte_rand <- rasterize(ext_buffer, ext_buf_r)
thatte_rand[] <- rnorm(length(getValues(thatte_rand)), 
                      mean = median(thatte_vals),
                      sd = sd(thatte_vals))
thatte_rand[thatte_rand<thatte_min] <- thatte_min
thatte_rand[thatte_rand>thatte_max] <- thatte_max

thatte_ring <- thatte_rand %>%
  crop(ring_buf) %>%
  mask(ring_buf)
  
plot(thatte_ring)

thatte_res_nodeless <- terra::cover(thatte_ring, extend(thatte, thatte_ring))
plot(thatte_res_nodeless, main = "Thatte et al.")
```

## Yumnam
```{r}
yumnam <- stack_res[[5]]
yumnam_vals <- getValues(yumnam) %>% as_tibble() %>% drop_na()
yumnam_vals <- yumnam_vals$value
yumnam_max <- max(yumnam_vals)
yumnam_min <- min(yumnam_vals)

yumnam_rand <- rasterize(ext_buffer, ext_buf_r)
yumnam_rand[] <- rnorm(length(getValues(yumnam_rand)), 
                      mean = median(yumnam_vals),
                      sd = sd(yumnam_vals))
yumnam_rand[yumnam_rand<yumnam_min] <- yumnam_min
yumnam_rand[yumnam_rand>yumnam_max] <- yumnam_max

yumnam_ring <- yumnam_rand %>%
  crop(ring_buf) %>%
  mask(ring_buf)
  
plot(yumnam_ring)

yumnam_res_nodeless <- terra::cover(yumnam_ring, extend(yumnam, yumnam_ring))
plot(yumnam_res_nodeless, main = "Yumnam et al.")
```


## Exporting for Circuitscape
```{r}
stack_res_nodeless <- stack(dutta_res_nodeless, mondal_res_nodeless, 
                            reddy_res_nodeless, thatte_res_nodeless, yumnam_res_nodeless) %>%
  `names<-` (c("Dutta_res_nodeless", "Mondal_res_nodeless", "Reddy_res_nodeless", 
                 "Thatte_res_nodeless", "Yumnam_res_nodeless"))

plot(stack_res_nodeless)

writeRaster(stack_res_nodeless, file.path("analysis_final/nodeless/"),
            names(stack_res_nodeless), bylayer = T, format = "ascii", overwrite = T)
```


# CS Notes
  - All layers run in Circuitscape standalone GUI wth ring_nodes as node layer
  - Ring buffer values randomly filled with mean = median of layer, sd = sd of layer
  - Nodes left hot (not set to 0), cumulative pairwise curent flow option selected
  - Results imported below

# Importing CS results
```{r}
## Masking CS results outside of individual study borders
dutta_cur_ndls.raw <- raster("C:/CS_outputs/CI_tiger/nodeless/Dutta_cur_nodeless_cum_curmap.asc", 
                        crs = utm.44n)
mondal_cur_ndls.raw <- raster("C:/CS_outputs/CI_tiger/nodeless/Mondal_cur_nodeless_cum_curmap.asc", 
                         crs = utm.44n)
reddy_cur_ndls.raw <- raster("C:/CS_outputs/CI_tiger/nodeless/Reddy_cur_nodeless_cum_curmap.asc", 
                        crs = utm.44n)
thatte_cur_ndls.raw <- raster("C:/CS_outputs/CI_tiger/nodeless/Thatte_cur_nodeless_cum_curmap.asc", 
                         crs = utm.44n)
yumnam_cur_ndls.raw <- raster("C:/CS_outputs/CI_tiger/nodeless/Yumnam_cur_nodeless_cum_curmap.asc", 
                         crs = utm.44n)

stack_cur_ndls.raw <- stack(dutta_cur_ndls.raw, mondal_cur_ndls.raw, reddy_cur_ndls.raw, 
                       thatte_cur_ndls.raw, yumnam_cur_ndls.raw) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp) %>%
  `names<-` (str_replace(names(stack_res), "res", "cur_ndls"))

plot(stack_cur_ndls.raw)
hist(stack_cur_ndls.raw)
par(mfrow = c(1,1), mar = c(5,4,4,2))

# writeRaster(stack_cur_ndls.raw, file.path("analysis_final/nodeless/", names(stack_cur_ndls.raw)), bylayer = T, format = "GTiff", overwrite = T)


# Importing TIFF Rasters 
cur_ndls_list <- lapply(list.files("analysis_final/nodeless/", pattern = "_cur_ndls.tif", 
                              all.files = TRUE, full.names = TRUE, ), raster)

stack_cur_ndls <- stack(cur_ndls_list) %>%
  crop(tiger_lyrs_shp) %>%
  mask(tiger_lyrs_shp)
plot(stack_cur_ndls)
hist(stack_cur_ndls, xlab = "Current")
par(mfrow = c(1,1), mar = c(5,4,4,2))
```

## Scale/average
```{r}
stack_cur_ndls.scaled <- rasterRescale(stack_cur_ndls) %>%
  `names<-` (str_glue("{names(stack_cur_ndls)}_scaled"))
plot(stack_cur_ndls.scaled)
hist(stack_cur_ndls.scaled)
par(mfrow = c(1,1), mar = c(5,4,4,2))

cur_ndls.scaled_avg <- mean(stack_cur_ndls.scaled)
plot(cur_ndls.scaled_avg, main = "Average Current (Scaled)")
hist(cur_ndls.scaled_avg, main = "Scaled Average current", xlab = "current", breaks = 30, xlim = c(0,0.5))

cur_ndls.scaled_cv <- calc(stack_cur_ndls.scaled, cv)
plot(cur_ndls.scaled_cv, main = "CV of Avg current (Scaled)")
hist(cur_ndls.scaled_cv, main = "Scaled CV current")

## Exporting files for GIS
# writeRaster(cur_ndls.scaled_avg, "analysis_final/nodeless/cur_ndls_scaled_avg.tif", overwrite = TRUE)
# writeRaster(cur_ndls.scaled_cv, "analysis_final/nodeless/cur_ndls_scaled_cv.tif", overwrite = TRUE)

## Import
cur_ndls.scaled_avg <- raster("analysis_final/cur_ndls/cur_ndls_scaled_avg.tif")
cur_ndls.scaled_cv <- raster("analysis_final/cur_ndls/cur_ndls_scaled_cv.tif")
```

### 20%ile CV
```{r}
# Low 20%ile CV areas
cur_ndls.cv_quant5 <- quantile(cur_ndls.scaled_cv, probs = seq(0,1,.2))
cur_ndls.scaled_low.cv20 <- cur_ndls.scaled_cv
cur_ndls.scaled_low.cv20[cur_ndls.scaled_low.cv20 > cur_ndls.cv_quant5[2]] <- NA
plot(cur_ndls.scaled_low.cv20)

cur_ndls.scaled_avg_low.cv20 <- mask(cur_ndls.scaled_avg, cur_ndls.scaled_low.cv20, maskvalue = NA)
plot(cur_ndls.scaled_avg_low.cv20)
# writeRaster(cur_ndls.scaled_avg_low.cv20, "analysis_final/nodeless/cur_ndls_scaled_avg_low_cv20.tif")

# Low 20%ile CV High Current 

## >50%ile of current masked with <20%ile CV
cur.avg_quant10 <- quantile(cur.scaled_avg, probs = seq(0,1,.1))
cur.scaled_avg_high.cur50 <- cur.scaled_avg
cur.scaled_avg_high.cur50[cur.scaled_avg_high.cur50 < cur.avg_quant10[6]] <- NA
plot(cur.scaled_avg_high.cur50)
cur.scaled_avg_high.cur50_low.cv20 <- mask(cur.scaled_avg_high.cur50, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur50_low.cv20)

# writeRaster(cur.scaled_avg_high.cur50, "analysis_final/cur/cur_scaled_avg_high_cur50.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur50_low.cv20, 
#             "analysis_final/cur/cur_scaled_avg_high_cur50_low_cv20.tif")

## >60%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur60 <- cur.scaled_avg
cur.scaled_avg_high.cur60[cur.scaled_avg_high.cur60 < cur.avg_quant10[7]] <- NA
plot(cur.scaled_avg_high.cur60)
cur.scaled_avg_high.cur60_low.cv20 <- mask(cur.scaled_avg_high.cur60, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur60_low.cv20)

# writeRaster(cur.scaled_avg_high.cur60, "analysis_final/cur/cur_scaled_avg_high_cur60.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur60_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur60_low_cv20.tif")

## >70%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur70 <- cur.scaled_avg
cur.scaled_avg_high.cur70[cur.scaled_avg_high.cur70 < cur.avg_quant10[8]] <- NA
plot(cur.scaled_avg_high.cur70)
cur.scaled_avg_high.cur70_low.cv20 <- mask(cur.scaled_avg_high.cur70, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur70_low.cv20)

# writeRaster(cur.scaled_avg_high.cur70, "analysis_final/cur/cur_scaled_avg_high_cur70.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur70_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur70_low_cv20.tif")

## >80%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur80 <- cur.scaled_avg
cur.scaled_avg_high.cur80[cur.scaled_avg_high.cur80 < cur.avg_quant10[9]] <- NA
plot(cur.scaled_avg_high.cur80)
cur.scaled_avg_high.cur80_low.cv20 <- mask(cur.scaled_avg_high.cur80, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur80_low.cv20)

# writeRaster(cur.scaled_avg_high.cur80, "analysis_final/cur/cur_scaled_avg_high_cur80.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur80_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur80_low_cv20.tif")

## >90%ile of current masked with <20%ile CV
cur.scaled_avg_high.cur90 <- cur.scaled_avg
cur.scaled_avg_high.cur90[cur.scaled_avg_high.cur90 < cur.avg_quant10[10]] <- NA
plot(cur.scaled_avg_high.cur90)
cur.scaled_avg_high.cur90_low.cv20 <- mask(cur.scaled_avg_high.cur90, cur.scaled_low.cv20, maskvalue = NA)
plot(cur.scaled_avg_high.cur90_low.cv20)

# writeRaster(cur.scaled_avg_high.cur90, "analysis_final/cur/cur_scaled_avg_high_cur90.tif", overwrite = T)
# writeRaster(cur.scaled_avg_high.cur90_low.cv20,
#             "analysis_final/cur/cur_scaled_avg_high_cur90_low_cv20.tif")
```

### 50%ile CV
```{r}
# Low 50%ile CV areas
cur.cv_quant2 <- quantile(cur.scaled_cv, probs = seq(0,1,.5))
cur.scaled_low.cv50 <- cur.scaled_cv
cur.scaled_low.cv50[cur.scaled_low.cv50 > cur.cv_quant2[2]] <- NA
plot(cur.scaled_low.cv50)

cur.scaled_avg_low.cv50 <- mask(cur.scaled_avg, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_low.cv50)
# writeRaster(cur.scaled_avg_low.cv50, "analysis_final/cur/cur_scaled_avg_low_cv50.tif")

## >60%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur60_low.cv50 <- mask(cur.scaled_avg_high.cur60, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur60_low.cv50)
# writeRaster(cur.scaled_avg_high.cur60_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur60_low_cv50.tif")

## >70%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur70_low.cv50 <- mask(cur.scaled_avg_high.cur70, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur70_low.cv50)
# writeRaster(cur.scaled_avg_high.cur70_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur70_low_cv50.tif")

## >80%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur80_low.cv50 <- mask(cur.scaled_avg_high.cur80, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur80_low.cv50)
# writeRaster(cur.scaled_avg_high.cur80_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur80_low_cv50.tif")

## >90%ile of current masked with <50%ile CV
cur.scaled_avg_high.cur90_low.cv50 <- mask(cur.scaled_avg_high.cur90, cur.scaled_low.cv50, maskvalue = NA)
plot(cur.scaled_avg_high.cur90_low.cv50)
# writeRaster(cur.scaled_avg_high.cur90_low.cv50,
#             "analysis_final/cur/cur_scaled_avg_high_cur90_low_cv50.tif")

```

# Quantitative Analysis

## Total CCA area (in sq km)

```{r}
cca_area <- as.numeric(length(cur.scaled_avg_high.cur70_low.cv20[!is.na(cur.scaled_avg_high.cur70_low.cv20)])) * 250*250/1e6
```


## High current/high agreement overlap
```{r}
#20th %ile CV used for high agreement
plot(cur.scaled_low.cv20)

#70th %ile current
low.cv20_high.cur70_mask <- mask(cur.scaled_low.cv20, cur.scaled_avg_high.cur70)
plot(low.cv20_high.cur70_mask)

length(low.cv20_high.cur70_mask[!is.na(low.cv20_high.cur70_mask)])/
  length(cur.scaled_low.cv20[!is.na(cur.scaled_low.cv20)])


# 80th %ile current
low.cv20_high.cur80_mask <- mask(cur.scaled_low.cv20, cur.scaled_avg_high.cur80)
plot(low.cv20_high.cur80_mask)

length(low.cv20_high.cur80_mask[!is.na(low.cv20_high.cur80_mask)])/
  length(cur.scaled_low.cv20[!is.na(cur.scaled_low.cv20)])

plot(cur.scaled_avg_high.cur80)
high.cur80_low.cv20_mask <- mask(cur.scaled_avg_high.cur80, cur.scaled_low.cv20)
plot(high.cur80_low.cv20_mask)

length(high.cur80_low.cv20_mask[!is.na(high.cur80_low.cv20_mask)])/
  length(cur.scaled_avg_high.cur80[!is.na(cur.scaled_avg_high.cur80)])


int_low.cv20_high.cur80 <- hull_low.cv20 %>%
  crop(hull_high.cur80) %>%
  mask(hull_high.cur80)
plot(int_low.cv20_high.cur80)

int_high.cur80_low.cv20 <- hull_high.cur80 %>%
  crop(hull_low.cv20) %>%
  mask(hull_low.cv20)
plot(int_high.cur80_low.cv20)

length(int_low.cv20_high.cur80[!is.na(int_low.cv20_high.cur80)])/
  length(hull_low.cv20[!is.na(hull_low.cv20)])

length(int_low.cv20_high.cur80[!is.na(int_low.cv20_high.cur80)])/
  length(hull_low.cv20[!is.na(hull_low.cv20)])

```
